---
title: "silt mica relation"
author: "Greta Dobetsberger"
date: '2023-10-19'
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Silt content and mica content relation

If we think about soil horizons high in silt and its effect on root development and root depth, it is also crucial to have a look into the relation between silt and mica content. This allows us to conclude 

```{r packages and data}
install.packages("GoodmanKruskal")
library(GoodmanKruskal)
library(readxl)
library(tidyverse)
library(ggplot2)
library(corrplot)
library(FactoMineR) # Pour l'analyse en composantes principales
library(factoextra)
library(psych)
install.packages("hrbrthemes")
library(hrbrthemes)
install.packages("gdtools")
library(gdtools)
library(viridis)
library(wesanderson)
install.packages("viridisLite")
dataPitF1 <- read_excel("FOSSE 1-4.xlsx", sheet=1, na="NA")
dataPitF2 <- read_excel("FOSSE 1-4.xlsx", sheet=2, na="NA")
dataPitF3 <- read_excel("FOSSE 1-4.xlsx", sheet=3, na="NA")
dataPitF4 <- read_excel("FOSSE 1-4.xlsx", sheet=4, na="NA")

dataPitF1$pit <- rep("F1", 100); dataPitF2$pit <- rep("F2", 100); dataPitF3$pit <- rep("F3", 100); dataPitF4$pit <- rep("F4", 100) #add a the variable pit to distinguish each sheet data
dataPitAll <- rbind(dataPitF1, dataPitF2, dataPitF3, dataPitF4)
```
Soil texture classes are difficult to bring in order according to their silt content. Soil horizons put into the LSA class for example can have more silt content than LAS horizons, but it can also contain the same proportion in silt or more. Hence, we grouped them into three soil texture classes.

group A: high silt content: LSA, LAS
group B: intermediate silt content: SL, SA
group C: low silt content: AS, A

```{r}
silt_content <- c(LSA= "high", LAS= "high", SL= "intermediate", SA="intermediate", AS="low", A="low")
dataPitAll$siltcontent<-silt_content[substr(dataPitAll$texture, 1, 3)]

#create new column with textures divided in three classes based upon their presence of silt
#to write it like Emilie, correct it like this: Auger$Silt_class <- ifelse(Auger$Texture_2023 %in% c("A", "Alo", "AS"), "low", ifelse(Auger$Texture_2023 %in% c("SL", "SA", "S"), "intermediate", "high"))
```
To visually explore the relationship between two ordinal/categorical variables and find out whether there can be a pattern found in our data, a mosaic plot was created.
Common correlation tests are not appropriate. 

```{r mosaicplot}
mosaicplot(table(dataPitAll$micas, dataPitAll$siltcontent), main= "Mica-Silt relationship in pits", color = TRUE, xlab="silt content", ylab="mica content", type="pearson")
#is type="pearson" really useful? unanswered question
```
To quantitatively test the association and assess the statistical significance of the association between the two ordinal variables, a statistical test like the chi-square test for independence was performed. It provides a p-value indicating the strength of the association.

```{r}
#Apparently the chi-square test is not appropriate cause it ignores the ordinal nature of the variables
#Perform a chi-square test for the Mica-Silt relationship for the pit data
print(chisq.test(dataPitAll$micas, dataPitAll$siltcontent))

print(fisher.test(dataPitAll$micas, dataPitAll$siltcontent)) # fisher's exact test due to small sampling size?


#Kruskal-Wallis rank sum test for ordinal data
kruskal_result <- kruskal.test(dataPitAll$micas, dataPitAll$siltcontent)
print(kruskal_result)


```

```{r}
#For testing a relationship between the two variables, which have both three categories now, different options have been trialed.

#cor.test(dataPitAll$micas,dataPitAll$siltcontent, method = 'spearman') doesnt work because of two ordinal variables
#correl<-cor_pb(dataPitAll$micas,dataPitAll$siltcontent) is an
# option for ordinal variables but Rstudio doesn't find the function although package is installed.

#another option suggested by chatgpt, but y axis in the wrong way (high mica content, low silt content)
#ggplot(dataPitAll, aes(x = siltcontent, fill = micas)) +
# geom_bar(position = "dodge") +
#  labs(title = "Clustered Bar Chart", x = "Silt content", y = "Mica content") +
#  scale_y_discrete(limits = rev(levels(dataPitAll$siltcontent)))
```
```




Now we are interested in the data from 2022, as there is more data points, hence, it looks more promising to find interesting information
```{r}
setwd("/Users/admin/Documents/GitHub/M2_EFT_projet_pedologie/Fichiers analyse")

dataAuger2022 <- read_excel("DataAuger2022.xlsx", sheet=1, na="NA")



#SCRIPT EMILIE

dataAuger2022 <- dataAuger2022 %>% rename_at('% pente', ~'pente')
dataAuger2022 <- dataAuger2022 %>% rename_at('Numéro horizon', ~'horizon')
dataAuger2022 <- dataAuger2022 %>% rename_at('profondeur limite inférieure (cm)2', ~'depth_inf')
dataAuger2022 <- dataAuger2022 %>% rename_at('profondeur limite supérieure (cm)', ~'depth_sup')
dataAuger2022 <- dataAuger2022 %>% rename_at('Epaisseur (cm)', ~'epaisseur')
dataAuger2022 <- dataAuger2022 %>% rename_at('Texture corrigée 2023', ~'Texture_2023')  #idk why it doesn't work here
dataAuger2022 <- dataAuger2022 %>% rename_at('EG %', ~'stone_elements')
dataAuger2022 <- dataAuger2022 %>% rename_at('Réserve eau utile (mm) NEW...35', ~'water_reservoir')

#create a new column for topography that only contains the number of the topography 
dataAuger2022$topo <- as.numeric(gsub("[^0-9]+", "", dataAuger2022$topographie))
#filter out all data that isn't on the slope (drainage type DLS) and make smaller usable dataframe
Auger_NA <- dataAuger2022 %>% 
  filter(topo %in% c("5", "4", "3")) %>%
  select(c("sondage_N", "Texture_2023", "depth_sup", "arbre_plus_proche", "topo", "mica")) 
##mica as well (different from Emilies df)
#take out NA variables
Auger <- Auger_NA %>% 
  na.omit   

#create new column with textures divided in three classes based upon their presence of silt
Auger$Silt_class <- ifelse(Auger$Texture_2023 %in% c("A", "Alo", "AS"), "low", ifelse(Auger$Texture_2023 %in% c("SL", "SA", "S"), "intermediate", "high"))

#silt 
```
From this part from the vegetation analysis (idk how we want to refer to emilies part correctly) we want to now analyse the relation between mica and silt content

```{r}
#mica content from 0-3 (4 categories)

#not sure whether mosaic plot is the right thing to do
##mosaicplot(table(Auger$Silt_class, Auger$mica), color = TRUE, xlab="silt content categories", ylab="mica content categories", type="pearson")

#two columns have to be transformed into a factor 
Auger$mica=as.factor(Auger$mica)
Auger$Silt_class=as.factor(Auger$Silt_class)

# Perform the chi-squared test (doesn't take into account that the variables ar ordinal)
chi_sq_test <- chisq.test(Auger$mica, Auger$Silt_class)
print(chi_sq_test)

# Kruskal-Wallis test (takes ordinal nature of the variables into account)
kruskal_result <- kruskal.test(Auger$mica, Auger$Silt_class)
print(kruskal_result)


```

To visually show the relation we use a mosaic plot.
```{r}
# Create a mosaic plot
mosaicplot(Auger$mica ~ Auger$Silt_class, data = Auger,
           color = TRUE, main = "Mica Silt relation", xlab="Mica content", ylab="Silt class")


#not really looking like anything useful...?

## didnt work due to package associationPlot(Auger$mica, Auger$Silt_class, main = "Association Plot")



```
This method is used to test and visualize associations between ordinal variables. They do not measure the strength or direction of a correlation as you would with continuous variables.

```{r}
Greta_pente<-read.csv("/Users/admin/Documents/GitHub/M2_EFT_projet_pedologie/Qgis/Greta_including_pente.csv")

#remove NA objects
Greta_pente_without_na <- Greta_pente %>% 
  na.omit 
  
Greta_pente_without_na$Slope=as_factor(Greta_pente_without_na$Slope)
Greta_pente_without_na$depth_sup=as.numeric(Greta_pente_without_na$depth_sup)
str(Greta_pente_without_na$Slope)
str(Greta_pente_without_na$depth_sup)


library(ggplot2)

ggplot(Greta_pente_without_na, aes(x = Slope, y = depth_sup)) +
  geom_boxplot(fill = "lightblue", color = "blue") +
  labs(
    title = "Soil Depth by Hill Position",
    x = "Hill Position",
    y = "HSD"
  )
linear_model=lm(Greta_pente_without_na$depth_sup~Greta_pente_without_na$Slope)
plot(linear_model)
anova(linear_model)


#or: 
#install.packages("analyzer")
#library(analyzer)
#CQassociation(Greta_pente_without_na$depth_sup, Greta_pente_without_na$Slope, method= "non-parametric")

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
