---
title: "pedo_Auger_Analysis"
author: "EmilieSanvito"
date: "2023-10-05"
output: html_document
---
###Pedologie projet

##data organization

```{r libraries}
#to empty your work environment
rm(list=ls()) 


library(tidyverse)
library(readxl)
library(dplyr)

library(ggforce)   #look again why I used it
library(knitr)

```

```{r upload and structure data}
#import data from last year project (36 Auger in the P16)
dataAuger2022 <- read_excel("DataAuger2022.xlsx", na = "NA")  

#create a new column for topography that only contains the number of the topography 
dataAuger2022$topo <- as.numeric(gsub("[^0-9]+", "", dataAuger2022$topographie))
#6: bas fonds
#5: versant bas de pente
#4: versant mi-pente
#3: versant haut de pente
#2: ...?
#1: plateau

#import botany data
Botany <- read_excel("DataAuger2022.xlsx", na = "NA", sheet = "Data Giacomo")  

#remove NA objects
Botany <- Botany %>% 
  na.omit  

#make a more precise df with variables we will use
Botany <- Botany %>% 
  select(c("TreeID", "Family", "Genus", "species", "DBH", "Xutm", "Yutm"))

```

```{r rename titles in data}

dataAuger2022 <- dataAuger2022 %>% rename_at('% pente', ~'pente')
dataAuger2022 <- dataAuger2022 %>% rename_at('Numéro horizon', ~'horizon')
dataAuger2022 <- dataAuger2022 %>% rename_at('profondeur limite inférieure (cm)2', ~'depth_inf')
dataAuger2022 <- dataAuger2022 %>% rename_at('profondeur limite supérieure (cm)', ~'depth_sup')
dataAuger2022 <- dataAuger2022 %>% rename_at('Epaisseur (cm)', ~'epaisseur')
dataAuger2022 <- dataAuger2022 %>% rename_at('Texture corrigée 2023', ~'Texture_2023')
dataAuger2022 <- dataAuger2022 %>% rename_at('EG %', ~'stone_elements')
dataAuger2022 <- dataAuger2022 %>% rename_at('Réserve eau utile (mm) NEW...35', ~'water_reservoir')

```


```{r}
Auger <- dataAuger2022 %>%
  select(c("sondage_N", "arbre_plus_proche", "topo", "depth_sup", "Texture_2023")) %>%
  na.omit

#create new column with textures divided in three classes based upon their presence of silt
Auger $Silt_class <- ifelse(Auger$Texture_2023 %in% c("A", "Alo", "AS"), "low", ifelse(Auger$Texture_2023 %in% c("SL", "SA", "S"), "intermediate", "high"))

```

We have 36 Auger points, of which some do not have corrected texture analysis so we do not consider these (22) and of these 22 there are 17 that are on the slope. Of these 17 only 13 have high silt class present in their texture data.

Of the 22 texture available, there are 16 that have high texture present


```{r part for correlation test}
#create a cleaner dataframe with only necessary informations
Auger_cor <- dataAuger2022 %>% 
  select(c("sondage_N", "depth_sup", "texture", "Texture_2023", "topo", "arbre_plus_proche"))

#create new texture variable that contains some non-corrected textures
Auger_cor$final_texture <- ifelse(
  is.na(dataAuger2022$Texture_2023) & !(dataAuger2022$sondage_N %in% c("15.1", "15.2", "15.5", "15.7", "19.5", "19.7", "20.5")),
  dataAuger2022$texture,
  dataAuger2022$Texture_2023
)

#create dataframe with only final texture as texture, and then only eliminate the rows for which we have not data of
Auger_cor <- Auger_cor %>%
  select(c("sondage_N", "depth_sup", "final_texture", "topo", "arbre_plus_proche")) %>%
  na.omit

#create new column with textures divided in three classes based upon their presence of silt
Auger_cor$Silt_class <- ifelse(Auger_cor$final_texture %in% c("A", "Alo", "AS"), "low", ifelse(Auger_cor$final_texture %in% c("SL", "SA", "S"), "intermediate", "high"))

#creata dataframe in which for every Auger point you have defined depth at which silt concentration becomes high, or if intermediate that it is 120 (!!Important to check this)
Auger_cor <- Auger_cor %>%
  group_by(sondage_N) %>%
  mutate(
    final_depth = ifelse(any(Silt_class == "high"), 
                         min(depth_sup[Silt_class == "high"]), 
                         120)
  ) %>%
  ungroup() %>%
  distinct(sondage_N, .keep_all = TRUE) %>%
  select(sondage_N, final_depth, topo, arbre_plus_proche)

#import elevation data of topography
Elevation <- read.csv("Elevation_exact.csv")

#merge two dataframes 
Elevation <- Elevation %>%
  select(c("sondage_N", "Elevation_Paracou_depth1")) %>%
  merge(Auger_cor, by.x = 'sondage_N', by.y = 'sondage_N')

#correlation test between topography and high silt depth appearance, spearman method instead of pearson due to non-normality (of final depth because much more values )
cor(Elevation$Elevation_Paracou_depth1, Elevation$final_depth, method = "spearman")

cor.test(Elevation$Elevation_Paracou_depth1, Elevation$final_depth, method = 
c("spearman"))

```

```{r little data visualisation for correlation}
hist(Elevation$final_depth)
hist(Elevation$Elevation_Paracou_depth1)
plot(Elevation$Elevation_Paracou_depth1, Elevation$final_depth)

ggplot(Elevation, aes(x=as.factor(topo), y=final_depth)) + geom_boxplot() + geom_point()

ggplot(Elevation, aes(x=Elevation_Paracou_depth1, y=final_depth)) + geom_point()
```

Assumptions for correlation test:
linearity, normality


Make a new column in which the textures are grouped based on the importance of silt present:

```{r cleaning data}

#filter out all data that isn't on a certain topography and make smaller usable dataframe
Auger <- dataAuger2022 %>% 
  filter(topo %in% c("3")) %>%
  select(c("sondage_N", "Texture_2023", "depth_sup", "arbre_plus_proche", "topo")) 

#take out NA variables
Auger <- Auger %>% 
  na.omit   

#create new column with textures divided in three classes based upon their presence of silt
Auger$Silt_class <- ifelse(Auger$Texture_2023 %in% c("A", "Alo", "AS"), "low", ifelse(Auger$Texture_2023 %in% c("SL", "SA", "S"), "intermediate", "high"))

```

!There are multiple DBH measurements because some trees have multiple stems!


```{r define silt appearance depth}

#creata dataframe in which for every Auger point you have defined depth at which silt concentration becomes high
Auger <- Auger %>% 
  group_by(sondage_N) %>%             #group by Auger
  filter(Silt_class == "high") %>%    #get only Augers for which the silt concentration is high
  arrange(sondage_N, depth_sup) %>%   #arrange the data so that every measurement in one Auger is arranged with depth
  distinct(sondage_N, .keep_all = TRUE) %>%  #keep only one value per Auger (the first depth)
  select(sondage_N,depth_sup, arbre_plus_proche, topo) %>% #maintain only three variables from the dataframe
  merge(Botany, by.x = 'arbre_plus_proche', by.y = 'TreeID' )   #merge with botany dataframe for common values of treeID and arbre plus proche

```

Points are grouped by distance from center point (arbre_plus_proche)

```{r create good buffer dataframe}

#write_csv(Botany, "Botany_export.csv")      
#export data to QGis

#upload data that contains zones of 10m around center Auger points
Buffer_points <- read.csv("Final_intersection.csv")

AGB_buffer <- Buffer_points %>%
  group_by(sondage_N) %>%
  summarise(amount_trees = n(),
            sp_richness = n_distinct(species)
            )

Final <- Auger %>%
  select(sondage_N, depth_sup, topo) %>%
  merge(AGB_buffer, by.x = 'sondage_N', by.y = 'sondage_N' )

```

I will make a new column that includes all textures, no NA anymore, so to keep all Augers
```{r ugh pause on this part}

#creates a new column for all textures combining the corrected and the not corrected ones (if topography is 5 or 6 and if it is not three points decided by visual that they are not good to)
dataAuger2022$final_texture <- ifelse(
  is.na(dataAuger2022$Texture_2023) & (dataAuger2022$topo == 5 | dataAuger2022$topo == 6) & !(dataAuger2022$sondage_N %in% c("15.1", "15.5", "15.7")),
  dataAuger2022$texture,
  dataAuger2022$Texture_2023
)


Greta1 <- dataAuger2022 %>% 
  select(c("sondage_N", "Texture_2023", "final_texture", "texture", "depth_sup", "arbre_plus_proche", "topo"))

#create new column with textures divided in three classes based upon their presence of silt
Greta1$Silt_class <- ifelse(Greta1$final_texture %in% c("A", "Alo", "AS"), "low", ifelse(Greta1$final_texture %in% c("SL", "SA", "S"), "intermediate", ifelse(Greta1$final_texture == "NA", "NA", "high")))

Greta_merged <- merge(Greta1, Botany, by.x = "arbre_plus_proche", by.y = "TreeID") %>%
  group_by(sondage_N) %>%             # Regroup by "sondage_N"
  filter(DBH == max(DBH)) %>%         # Filter to keep rows with the highest DBH
  distinct(sondage_N, .keep_all = TRUE) # Keep only one row per sondage_N

write_csv(Greta_merged, "Greta.csv")   #for greta

Greta <- Greta1 %>%
  group_by(sondage_N) %>%             # Group by Auger
  filter(Silt_class == "high") %>%    # Get only Augers with high silt concentration
  arrange(sondage_N) %>%              # Arrange data by "sondage_N"
  left_join(Botany, by = c("arbre_plus_proche" = "TreeID")) %>% # Left join with Botany
  group_by(sondage_N) %>%             # Regroup by "sondage_N"
  filter(DBH == max(DBH)) %>%         # Filter to keep rows with the highest DBH
  distinct(sondage_N, .keep_all = TRUE) %>% # Keep only one row per sondage_N
  select(sondage_N, depth_sup, arbre_plus_proche, topo, DBH, Xutm, Yutm) # Maintain selected variables


```


For map in QGis we exported a dataframe with X and Y coordinates and auger points, with every its designed depth.

## data visualisation

```{r quick plots}

plot(Final_df$depth_sup, Final_df$amount_trees)
plot(Final_df$depth_sup, Final_df$sp_richness)
plot(Final_df$depth_sup, Final_df$DBH)
```

## test if there is an effect of the depth of the silt horizon appearance on the number of trees


check correlation between depth and topo!!
make boxplot of profondeur de limon with topography, be aware of this

if there is a correlation between the silt depth and distance of bas fonds => will complicate your things because we know there is a link between topo and species so then you cannot say anything 

it might be interesting to look at type1 test because then you can put importance to your first variables that you put in

