---
title: "pedo_Auger_Analsysis"
author: "EmilieSanvito"
date: "2023-10-05"
output: html_document
---
###Pedologie projet

##data organization

```{r libraries}
rm(list=ls()) #to empty your work environment


library(tidyverse)
library(readxl)
library(dplyr)

#install.packages("ggforce")
library(ggforce)   #look again why I used it
library(knitr)

library(geosphere)  #library for calculating distances of latitudes


#install.packages("memoiR")
#library(memoiR)

```

```{r upload and structure data}

dataAugur2022 <- read_excel("DataAuger2022.xlsx", na = "NA")   #import data from last year project (36 augur in the P16)

dataBotany2022_NA <- read_excel("DataAuger2022.xlsx", na = "NA", sheet = "Data Giacomo")  #import botany data
dataBotany2022 <- dataBotany2022_NA %>% 
  na.omit  #remove NA objects

```

```{r rename titles in data}

dataAugur2022 <- dataAugur2022 %>% rename_at('% pente', ~'pente')
dataAugur2022 <- dataAugur2022 %>% rename_at('Numéro horizon', ~'horizon')
dataAugur2022 <- dataAugur2022 %>% rename_at('profondeur limite inférieure (cm)2', ~'depth_inf')
dataAugur2022 <- dataAugur2022 %>% rename_at('profondeur limite supérieure (cm)', ~'depth_sup')
dataAugur2022 <- dataAugur2022 %>% rename_at('Epaisseur (cm)', ~'epaisseur')
dataAugur2022 <- dataAugur2022 %>% rename_at('Texture corrigée 2023', ~'Texture_2023')  #idk why it doesn't work here
dataAugur2022 <- dataAugur2022 %>% rename_at('EG %', ~'stone_elements')
dataAugur2022 <- dataAugur2022 %>% rename_at('Réserve eau utile (mm) NEW...35', ~'water_reservoir')

```


Make a new column in which the textures are grouped based on the importance of silt present:

```{r}
Augur_NA <- dataAugur2022 %>% filter(Classification == "Drainage latéral superficiel") %>%
  select(c("sondage_N", "Texture_2023", "depth_sup", "arbre_plus_proche")) 
 #filter out all data that isn't on the slope (drainage type DLS) and make smaller usable dataframe

Augur <- Augur_NA %>% 
  na.omit   #take out NA variables

Augur$Silt_class <- ifelse(Augur$Texture_2023 %in% c("A", "Alo", "AS", "SA", "S"), "low", ifelse(Augur$Texture_2023 == "SL", "intermediate", "high"))  
#create new column with textures divided in three classes based upon their presence of silt
#class A has most amount of silt, class C has least amount of silt

Augur$Silt <- ifelse(Augur$Texture_2023 %in% c("A", "Alo", "AS", "SA", "SL", "S"), "low", "high")

Botany <- dataBotany2022 %>% 
  select(c("TreeID", "Family", "Genus", "species", "Xutm", "Yutm")) #make a more precise df with variables we will use

```

## Data visualizations


```{r define silt appearance depth}

df_restructured <- Augur %>% 
  group_by(sondage_N) %>% 
  filter(Texture_simplified == "high") %>% 
  arrange(sondage_N, depth_sup) %>% 
  distinct(sondage_N, .keep_all = TRUE) %>% 
  select(sondage_N,depth_sup, arbre_plus_proche) %>% 
  merge(Botany, by.x = 'arbre_plus_proche', by.y = 'TreeID' )

# remade a df in which for each augur measurement the depth at which the silt occurs in high quantities (so class A) is noted down
# last merge: create a merged dataframe of both the botany and the augur data based on the columns "arbre_plus_proche" and "TreeID" ; this time in which the data is not repeated for all the horizons

df_restructured <- Augur %>% 
  group_by(sondage_N) %>% 
  arrange(sondage_N, depth_sup) %>%
  


df_restructured$sondage_N <- as.numeric(df_merged_restructured$sondage_N)

```

Create dataframe where points are grouped by distance from center point (arbre_plus_proche)

```{r create good buffer dataframe}

#write_csv(Botany, "Botany_export.csv")      # export data to QGis
#write_csv(df_merged_restructured, "Augur_points.csv") # export data to bring to QGis


Buffer_points <- read.csv("buffer_points.csv")  #upload data that contains zones of 5m around center augur points


Final_df <- Buffer_points %>%
  group_by(sondage_N) %>%
  summarise(amount_trees = n(),
            sp_richness = n_distinct(species) 
            )
  
  
result <- Buffer_points %>%
  group_by(sondage_N, Family) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  slice(1) %>%
  ungroup() %>%
  group_by(sondage_N) %>%
  summarise(most_abundant_family = Family[1], 
            most_abundant_count = count[1],
            total_families_seen = n_distinct(Family))


```

## test if there is an effect of the depth of the silt horizon appearance on the number of trees





## extra

```{r code to find package of a function}
#install.packages("sos")  #package for function find Fn,
library(sos)
findFn("geosphere")  #function to learn which package is necessary for your function
```



Idea now is to check link between texture_class and the aboveground biomass


